---
title: "Behavior Association Rules - Daily"
author: "Brad Cannell"
date: "June 8, 2016"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

# Set working directory
knitr::opts_knit$set(root.dir = "/Users/bradcannell/Dropbox/Research/mChat")
```

-------------------------------------------------------------------------------

**Load packages**
```{r message=FALSE}
# CRAN packages
library(dplyr)
library(ggplot2)

# devtools::install_github("mbcann01/dataclean")
library(dataclean)
```

**Load data**  
```{r}
load("data/daily_april_2016.RData")

# Sort by case number and date
daily <- dplyr::arrange(daily, case_number, date)
about_data(daily)
```

-------------------------------------------------------------------------------   

&nbsp;

## Introduction

We are interested in understanding which behaviors, if any, tend to occur together. In order to explore patterns of co-occurrence we use association rules analysis (Lantz, 2013; Miner, Nisbet, & Elder IV, 2009). 

Candidate behaviors for inclusion in the analysis include:

1. Physical activity

2. Diet

3. Medication use

4. Free-time activities

5. Meaningful interactions

6. Substance use

In addition to behaviors, we may also explore the co-occurrence of mood states.


**References:**

Lantz, B. (2013). Machine learning with R. Packt Publishing Ltd.
Miner, G., Nisbet, R., & Elder IV, J. (2009). Handbook of statistical analysis and data mining applications. Academic Press.

&nbsp;

-------------------------------------------------------------------------------   

&nbsp;

Before moving on with this analysis, I need to do a basic exploritory analysis of all the individual behaviors.

I will also need to dichotomize many of the behavioral outcomes for the association rules analysis.


&nbsp;

-------------------------------------------------------------------------------   

&nbsp;

#### Session Info:
```{r echo=FALSE}
sessionInfo()
```
