---
title: "Clean Daily EMA Data"
date: "July 22, 2016"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

# Set working directory
knitr::opts_knit$set(root.dir = "/Users/bradcannell/Dropbox/Research/mChat")
```

-------------------------------------------------------------------------------

# Load packages and functions

```{r load_packages, message=FALSE}
# CRAN packages
library(readxl)
library(dplyr)
library(tidyr)
library(readr)
library(testthat)

# devtools::install_github("mbcann01/dataclean")
library(dataclean) 
# Use this instead of str() to prevent showing raw data values
```



# Import data   

1. Create character vector of column names and types (necessary because of sparsely populated columns).   
2. Import data assigning names and types.   
3. Remove any empty rows that were imported erroneously.   
4. Test for expected number of variables.   

```{r import_data}
col_info <- c( 
  # "col_name",           "col_type"
  "case_number",          "text",
  "date",                 "date",
  "touch_name",           "text",
  "happy",                "text",
  "frustrated",           "text",
  "sad",                  "text",
  "worried",              "text",
  "restless",             "text",
  "excited",              "text",
  "calm",                 "text",
  "lonely",               "text",
  "bored",                "text",
  "sluggish",             "text",
  "pre_location",         "text",
  "pre_action",           "text",
  "pre_who1",             "text",
  "pre_who2",             "numeric",
  "pre_yesterday",        "text",
  "min_walk",             "text",
  "min_walk_c",           "numeric",
  "min_run",              "text",
  "min_run_c",            "numeric",
  "min_act",              "text",
  "min_act_c",            "numeric",
  "min_sit",              "text",
  "min_sit_c",            "numeric",
  "pedometer1",           "text",
  "pedometer2",           "numeric",
  "steps",                "numeric",
  "phys_act_sat1",        "text",
  "phys_act_sat2",        "numeric",
  "serv_fruit1",          "text",
  "serv_fruit2",          "numeric",
  "serv_veg",             "text",
  "serv_ssb1",            "text",
  "serv_ssb2",            "numeric",
  "serv_sweets1",         "text",
  "serv_sweets2",         "numeric",
  "serv_meat",            "text",
  "food_sat1",            "text",
  "food_sat2",            "numeric",
  "take_meds1",           "text",
  "take_meds2",           "numeric",
  "which_meds1",          "text",
  "why_no_meds",          "text",
  "free_time",            "text",
  "free_time_act",        "text",
  "free_time_sat1",       "text",
  "free_time_sat2",       "numeric",
  "min_talk1",            "text",
  "min_talk2",            "numeric",
  "min_group1",           "text",
  "min_group2",           "numeric",
  "interaction_sat1",     "text",
  "interaction_sat2",     "numeric",
  "subs_check1",          "text",
  "subs_check2",          "numeric",
  "num_drinks1",          "text",
  "num_drinks2",          "numeric",
  "subs_time",            "text",
  "first_drink",          "text",
  "last_drink",           "text",
  "num_cigs1",            "text",
  "num_cigs2",            "numeric",
  "first_cig",            "text",
  "last_cig",             "text",
  "opiate",               "text",
  "num_heroin",           "text",
  "num_opiate",           "text",
  "stim",                 "text",
  "num_coc",              "text",
  "num_stim",             "text",
  "case_number_2",        "blank",
  "alert_id",             "blank",
  "alert_time",           "blank",
  "alert_type",           "blank",
  "initial_alert_time",   "text",
  "initial_alert_action", "blank",
  "start_time",           "text",
  "end_time",             "text",
  "completed_1",          "blank",
  "completed_2",          "blank",
  "gps",                  "text",
  "order_test",           "blank",
  "which_meds2",          "text",
  "dob",                  "text",
  "gender",               "text",
  "hispanic",             "text",
  "race",                 "text",
  "race_other",           "text"
)

# Import data
daily <- read_excel(
  path = "data/raw data/All EMA Data 7_19_16.xls",
  sheet = "Daily EMA",
  col_names = col_info[seq(1, length(col_info), 2)],
  col_types = col_info[seq(2, length(col_info), 2)],
  skip = 1)
rm(col_info)

# Convert selected variables to type numeric
daily <- daily %>%
  mutate(
    num_coc  = as.integer(num_coc),
    num_stim = as.integer(num_stim)
  )

# Remove blank rows
index <- all_na(daily)
if (length(index) == 0) {
  cat("Now rows deleted because of missing values for all variables")
} else if (index > 0) {
  daily <- daily[-index, ]
  cat(paste("Row(s)", index, "deleted because all values were missing")) 
}
rm(index)

# Data check
about_data(daily)
# 2016-07-22: 7,378 observations and 82 variables in the data

test_that("The number of variables hasn't changed", {
  expect_equal(ncol(daily), 82)
})

test_that("Every row contains a case_number", {
  expect_equal(sum(is.na(daily$case_number)), 0)
})
```

-------------------------------------------------------------------------------   

&nbsp; 

# Convert date / time variables   

```{r convert_date, warning=FALSE}
# Convert initial_alert_time, start_time, and end_time to class POSIXct
daily <- daily %>%
  # Separate initial alert time into hour and part of hour columns (x, y)
  separate(
    col = initial_alert_time, 
    into = c("x", "y"), 
    remove = FALSE) %>% 
  
  # Convert initial_alert_time to the form hours:minutes
  # Convert start_time and end_time to class POSIX
  mutate(
    x = as.integer(x),
    y = as.integer(y),
    y = y / 100,
    # Covert to minutes
    y = round(y * 60),
    # If initial alert time wasn't missing, set y = 0
    y = if_else(!is.na(initial_alert_time) & is.na(y), 0, y),
    
    # Create new POSIX class variable by combining date and time
    initial_alert_time = as.POSIXct(paste(date, x, y), format = "%Y-%m-%d %H %M"),
    start_time = as.POSIXct(paste(date, start_time), format = "%Y-%m-%d %H:%M"),
    end_time = as.POSIXct(paste(date, end_time), format = "%Y-%m-%d %H:%M"),
    
    # Drop x and y
    x = NULL,
    y = NULL
  )
```

-------------------------------------------------------------------------------   

&nbsp;  

# Create dummy variables for multiple answer questions

```{r dummy_coding}
daily <- daily %>%
  mutate(
    # Where are you right now?
    loc_bus        = ifelse(grepl("Bus", pre_location), 1, 0),
    loc_church     = ifelse(grepl("Church", pre_location), 1, 0),
    loc_home       = ifelse(grepl("Home", pre_location), 1, 0),
    loc_frifam     = ifelse(grepl("Friend's", pre_location), 1, 0),
    loc_store      = ifelse(grepl("Store", pre_location), 1, 0),
    loc_outdoors   = ifelse(grepl("Outdoors", pre_location), 1, 0),
    loc_other      = ifelse(grepl("Other", pre_location), 1, 0),
    loc_restaurant = ifelse(grepl("Restaurant", pre_location), 1, 0),
    loc_school     = ifelse(grepl("school", pre_location), 1, 0),
    loc_work       = ifelse(grepl("Work", pre_location), 1, 0),

    # Data check
    # View(select(test, pre_location, loc_bus:loc_work))
    
    # What were you doing right before your phone rang/vibrated?
    act_other      = ifelse(grepl("Other", pre_action), 1, 0),
    act_sit        = ifelse(grepl("Sitting", pre_action), 1, 0),
    act_sleep      = ifelse(grepl("Sleeping", pre_action), 1, 0),
    act_stand      = ifelse(grepl("Standing", pre_action), 1, 0),
    act_walk       = ifelse(grepl("Walking", pre_action), 1, 0),
    act_talk       = ifelse(grepl("Talking", pre_action), 1, 0),

    # Data check
    # View(select(test, pre_action, act_other:act_talk))
    
    # Which of these things did you do yesterday?
    yest_bike      = ifelse(grepl("Walk or bicycle", pre_yesterday), 1, 0),
    yest_run       = ifelse(grepl("running", pre_yesterday), 1, 0),
    yest_cleaning  = ifelse(grepl("cleaning", pre_yesterday), 1, 0),
    yest_none      = ifelse(grepl("None", pre_yesterday), 1, 0),

    # Data check
    # View(select(test, pre_yesterday, yest_bike:yest_none))
    
    # Which medication didn’t you take? 
    medtype_dep    = ifelse(grepl("Depression", which_meds2), 1, 0),
    medtype_psy    = ifelse(grepl("Psychiatric", which_meds2), 1, 0),
    medtype_ast    = ifelse(grepl("Asthma", which_meds2), 1, 0),
    medtype_bp     = ifelse(grepl("Blood pressure", which_meds2), 1, 0),
    medtype_diab   = ifelse(grepl("Diabetes", which_meds2), 1, 0),
    medtype_pain   = ifelse(grepl("Pain", which_meds2), 1, 0),
    medtype_chol   = ifelse(grepl("Cholesterol", which_meds2), 1, 0),
    medtype_anti   = ifelse(grepl("Antibiotic", which_meds2), 1, 0),
    medtype_other  = ifelse(grepl("Other", which_meds2), 1, 0),

    # Data check
    # View(select(test, which_meds2, medtype_dep:medtype_other))
    
    # Why didn’t you take medications?
    meds_ran_out   = ifelse(grepl("Ran", why_no_meds), 1, 0),
    meds_forgot    = ifelse(grepl("Forgot", why_no_meds), 1, 0),
    meds_no_need   = ifelse(grepl("need", why_no_meds), 1, 0),
    meds_side      = ifelse(grepl("Side effects", why_no_meds), 1, 0),
    meds_lost      = ifelse(grepl("lost", why_no_meds), 1, 0),
    meds_other     = ifelse(grepl("Other", why_no_meds), 1, 0),

    # Data check
    # View(select(test, why_no_meds, meds_ran_out:meds_other))
    
    # What did you do in your free time yesterday?
    free_none      = ifelse(grepl("None", free_time_act), 1, 0),
    free_tv        = ifelse(grepl("Watched TV", free_time_act), 1, 0),
    free_games     = ifelse(grepl("computer games", free_time_act), 1, 0),
    free_movie     = ifelse(grepl("movie", free_time_act), 1, 0),
    free_internet  = ifelse(grepl("Surfed the i", free_time_act), 1, 0),
    free_book      = ifelse(grepl("Read a book", free_time_act), 1, 0),
    free_music     = ifelse(grepl("music or played", free_time_act), 1, 0),
    free_craft     = ifelse(grepl("Arts,", free_time_act), 1, 0),
    free_cleaned   = ifelse(grepl("Cleaned", free_time_act), 1, 0),
    free_pets      = ifelse(grepl("of pets", free_time_act), 1, 0),
    free_pray      = ifelse(grepl("Prayed,", free_time_act), 1, 0),
    free_sport     = ifelse(grepl("Played a sport", free_time_act), 1, 0),
    free_walk      = ifelse(grepl("or jog", free_time_act), 1, 0),
    free_cards     = ifelse(grepl("cards", free_time_act), 1, 0),
    free_shop      = ifelse(grepl("Went shopping", free_time_act), 1, 0),
    free_volunteer = ifelse(grepl("Volunteered", free_time_act), 1, 0),
    free_social    = ifelse(grepl("Socialized with", free_time_act), 1, 0),
    
    # Data check
    # View(select(test, free_time_act, free_none:free_social))
    
    # Which of the following did you use yesterday, even once?
    sub_alc        = ifelse(grepl("Alcohol", subs_check1), 1, 0),
    sub_cig        = ifelse(grepl("Tobacco", subs_check1), 1, 0),
    sub_mar        = ifelse(grepl("Marijuana", subs_check1), 1, 0),
    sub_opi        = ifelse(grepl("Opiates", subs_check1), 1, 0),
    sub_stim       = ifelse(grepl("Stimulants", subs_check1), 1, 0),
    sub_herb       = ifelse(grepl("Herbal", subs_check1), 1, 0),
    sub_other      = ifelse(grepl("Another drug", subs_check1), 1, 0),
    sub_none       = ifelse(grepl("None", subs_check1), 1, 0),
    
    # Data check
    # View(select(test, subs_check1, sub_alc:sub_none))
    
    # Which of the following opiates did you used yesterday?
    op_heroin      = ifelse(grepl("Heroin", opiate), 1, 0),
    op_methadone   = ifelse(grepl("Methadone", opiate), 1, 0),
    op_pills       = ifelse(grepl("Pills", opiate), 1, 0),
    op_other       = ifelse(grepl("Other", opiate), 1, 0),
    
    # Data check
    # View(select(test, opiate, op_heroin:op_other))
    
    # Which of the following stimulants did you used yesterday?
    stim_coc      = ifelse(grepl("Cocaine/Crack Cocaine", stim), 1, 0),
    stim_meth     = ifelse(grepl("Methamphetamines", stim), 1, 0),
    stim_pills    = ifelse(grepl("Pills", stim), 1, 0),
    stim_other    = ifelse(grepl("Other", stim), 1, 0)
    
    # Data check
    # View(filter(test, !is.na(stim)) %>% select(stim, stim_coc:stim_other))
  )

# Data check
about_data(daily)
# 2016-07-22: 7,378 observations and 150 variables in the data

test_that("The number of variables hasn't changed", {
  expect_equal(ncol(daily), 150)
})
```

-------------------------------------------------------------------------------   

&nbsp;  

# Dichotomize selected categorical variables

```{r dichotomize}
daily <- daily %>%
  mutate(
    
    heavy_drink = ifelse(num_drinks2 >= 5, 1, 0),
    
    any_fruit   = ifelse(serv_fruit1 > 0, 1, 0),
    any_veg     = ifelse(serv_veg > 0, 1, 0),
    any_ssb     = ifelse(serv_ssb1 > 0, 1, 0),
    any_sweets  = ifelse(serv_sweets1 > 0, 1, 0),
    any_meat    = ifelse(serv_meat > 0, 1, 0),
    
    any_talk    = ifelse(min_talk1 > "15 or less minutes", 1, 0),
    any_group   = ifelse(min_group1 > "15 or less minutes", 1, 0)
  )

# Data check
# View(daily[c("min_group1", "any_group")])
```

Note: There was no way for a participant to indicate that they had zero minutes of meaningful individual or group interaction on the previous day (`min_talk1` and `min_group1`). `any_talk` and `any_group` really represent at least 15 minutes of meaningful conversation or group interaction on the previous day.

-------------------------------------------------------------------------------   

&nbsp;

# Covert categorical variables to class factor  

```{r to_factor}
# Create factor labels
likert <- c(
  "Strongly disagree",
  "Disagree",
  "Neutral",
  "Agree",
  "Strongly agree"
)

yesno <- c(
  "No",
  "Yes"
)

minutes <- c(
  "10 or less minutes",
  "11-20",
  "21-30",
  "31-40",
  "41-50",
  "more than 50"
)

hours <- c(
  "4 or fewer hours",
  "More than 4 hours- 6 hours",
  "More than 6 hours- 8 hours",
  "More than 8 hours- 10 hours",
  "More than 10 hours - 12 hours",
  "More than 12 hours"
)

hours2 <- c(
  "2 or fewer hours",
  "More than 2 hours- 4 hours",
  "More than 4 hours- 6 hours",
  "More than 6 hours- 8 hours",
  "More than 8 hours- 10 hours",
  "More than 10 hours"
)

satisfy <- c(
  "Not at all",
  "A little",
  "Somewhat",
  "Very",
  "Extremely satisfied"
)

meaningful_time <- c(
  "15 or less minutes",
  "16-30 minutes",
  "31 min- 1 hour",
  "More than 1 hour- 2 hours",
  "More than 2 hours- 3 hours",
  "More than 3 hours- 4 hours",
  "More than 4 hours"
)

drinks <- c(
  "0 drinks",
  "1 drink",
  "2 drinks",
  "3 drinks",
  "4 drinks",
  "5 drinks",
  "6 drinks",
  "7 or more drinks"
)

time <- c(
  "Before 9 AM",
  "9:00 AM-11:59 AM", 
  "12:00-2:59 PM", 
  "3:00-5:59 PM",
  "6:00-8:59 PM", 
  "9 PM or later"
)

cig_count <- c(
  "0 cigarettes",
  "One puff to one cigarette",
  "2-5 cigarettes",
  "6-10 cigarettes",
  "11-15 cigarettes",
  "16-20 cigarettes",
  "More than 20 cigarettes"
)

cigs_time <- c(
  "within 5 minutes",
  "6-30 minutes", 
  "31-60 minutes", 
  "after 60 minutes"
)

how_many <- c(
  "0",
  "1", 
  "2", 
  "3", 
  "4", 
  "5 or more"
)

daily <- daily %>%
  mutate(
     
    happy           = factor(happy, levels = likert, labels = likert),
    frustrated      = factor(frustrated, levels = likert, labels = likert),
    sad             = factor(sad, levels = likert, labels = likert),
    worried         = factor(worried, levels = likert, labels = likert),
    restless        = factor(restless, levels = likert, labels = likert),
    excited         = factor(excited, levels = likert, labels = likert),
    calm            = factor(calm, levels = likert, labels = likert),
    lonely          = factor(lonely, levels = likert, labels = likert),
    bored           = factor(bored, levels = likert, labels = likert),
    sluggish        = factor(sluggish, levels = likert, labels = likert),
    
    loc_bus         = factor(loc_bus, levels = 0:1, labels = yesno),
    loc_church      = factor(loc_church, levels = 0:1, labels = yesno),
    loc_home        = factor(loc_home, levels = 0:1, labels = yesno),
    loc_frifam      = factor(loc_frifam, levels = 0:1, labels = yesno),
    loc_store       = factor(loc_store, levels = 0:1, labels = yesno),
    loc_outdoors    = factor(loc_outdoors, levels = 0:1, labels = yesno),
    loc_other       = factor(loc_other, levels = 0:1, labels = yesno),
    loc_restaurant  = factor(loc_restaurant, levels = 0:1, labels = yesno),
    loc_school      = factor(loc_school, levels = 0:1, labels = yesno),
    loc_work        = factor(loc_work, levels = 0:1, labels = yesno),
    
    act_other       = factor(act_other, levels = 0:1, labels = yesno),
    act_sit         = factor(act_sit, levels = 0:1, labels = yesno),
    act_sleep       = factor(act_sleep, levels = 0:1, labels = yesno),
    act_stand       = factor(act_stand, levels = 0:1, labels = yesno),
    act_walk        = factor(act_walk, levels = 0:1, labels = yesno),
    act_talk        = factor(act_talk, levels = 0:1, labels = yesno),
    
    pre_who         = factor(pre_who2, levels = 0:1, labels = yesno),
    
    yest_bike       = factor(yest_bike, levels = 0:1, labels = yesno),
    yest_run        = factor(yest_run, levels = 0:1, labels = yesno),
    yest_cleaning   = factor(yest_cleaning, levels = 0:1, labels = yesno),
    yest_none       = factor(yest_none, levels = 0:1, labels = yesno),
    min_walk        = factor(min_walk, levels = minutes, labels = minutes),
    min_run         = factor(min_run, levels = minutes, labels = minutes),
    min_act         = factor(min_act, levels = minutes, labels = minutes),
    min_sit         = factor(min_sit, levels = hours, labels = hours),
    pedometer       = factor(pedometer2, levels = 0:1, labels = yesno),
    phys_act_sat    = factor(phys_act_sat1, levels = satisfy, labels = satisfy),
    
    serv_fruit      = factor(serv_fruit1, levels = how_many, labels = how_many),
    serv_veg        = factor(serv_veg, levels = how_many, labels = how_many),
    serv_ssb        = factor(serv_ssb1, levels = how_many, labels = how_many),
    serv_sweets     = factor(serv_sweets1, levels = how_many, labels = how_many),
    serv_meat       = factor(serv_meat, levels = how_many, labels = how_many),
    food_sat        = factor(food_sat1, levels = satisfy, labels = satisfy),
    
    any_fruit       = factor(any_fruit, levels = 0:1, labels = yesno),
    any_veg         = factor(any_veg, levels = 0:1, labels = yesno),
    any_ssb         = factor(any_ssb, levels = 0:1, labels = yesno),
    any_sweets      = factor(any_sweets, levels = 0:1, labels = yesno),
    any_meat        = factor(any_meat, levels = 0:1, labels = yesno),
    
    take_meds       = factor(take_meds2, levels = 0:1, labels = yesno),
    
    medtype_dep     = factor(medtype_dep, levels = 0:1, labels = yesno),
    medtype_psy     = factor(medtype_psy, levels = 0:1, labels = yesno),
    medtype_ast     = factor(medtype_ast, levels = 0:1, labels = yesno),
    medtype_bp      = factor(medtype_bp, levels = 0:1, labels = yesno),
    medtype_diab    = factor(medtype_diab, levels = 0:1, labels = yesno),
    medtype_pain    = factor(medtype_pain, levels = 0:1, labels = yesno),
    medtype_chol    = factor(medtype_chol, levels = 0:1, labels = yesno),
    medtype_anti    = factor(medtype_anti, levels = 0:1, labels = yesno),
    medtype_other   = factor(medtype_other, levels = 0:1, labels = yesno),
    
    meds_ran_out    = factor(meds_ran_out, levels = 0:1, labels = yesno),
    meds_forgot     = factor(meds_forgot, levels = 0:1, labels = yesno),
    meds_no_need    = factor(meds_no_need, levels = 0:1, labels = yesno),
    meds_side       = factor(meds_side, levels = 0:1, labels = yesno),
    meds_lost       = factor(meds_lost, levels = 0:1, labels = yesno),
    meds_other      = factor(meds_other, levels = 0:1, labels = yesno),
    
    free_time       = factor(free_time, levels = hours2, labels = hours2),
    free_none       = factor(free_none, levels = 0:1, labels = yesno),
    free_tv         = factor(free_tv, levels = 0:1, labels = yesno),
    free_games      = factor(free_games, levels = 0:1, labels = yesno),
    free_movie      = factor(free_movie, levels = 0:1, labels = yesno),
    free_internet   = factor(free_internet, levels = 0:1, labels = yesno),
    free_book       = factor(free_book, levels = 0:1, labels = yesno),
    free_music      = factor(free_music, levels = 0:1, labels = yesno),
    free_craft      = factor(free_craft, levels = 0:1, labels = yesno),
    free_cleaned    = factor(free_cleaned, levels = 0:1, labels = yesno),
    free_pets       = factor(free_pets, levels = 0:1, labels = yesno),
    free_pray       = factor(free_pray, levels = 0:1, labels = yesno),
    free_sport      = factor(free_sport, levels = 0:1, labels = yesno),
    free_walk       = factor(free_walk, levels = 0:1, labels = yesno),
    free_cards      = factor(free_cards, levels = 0:1, labels = yesno),
    free_shop       = factor(free_shop, levels = 0:1, labels = yesno),
    free_volunteer  = factor(free_volunteer, levels = 0:1, labels = yesno),
    free_social     = factor(free_social, levels = 0:1, labels = yesno),
    free_time_sat   = factor(free_time_sat1, levels = satisfy, labels = satisfy),
    
    min_talk        = factor(min_talk1, levels = meaningful_time, labels = meaningful_time),
    min_group       = factor(min_group1, levels = meaningful_time, labels = meaningful_time),
    interaction_sat = factor(interaction_sat1, levels = satisfy, labels = satisfy),
    
    any_talk        = factor(any_talk, levels = 0:1, labels = yesno),
    any_group       = factor(any_group, levels = 0:1, labels = yesno),
    
    sub_alc         = factor(sub_alc, levels = 0:1, labels = yesno),
    heavy_drink     = factor(heavy_drink, levels = 0:1, labels = yesno),
    sub_cig         = factor(sub_cig, levels = 0:1, labels = yesno),
    sub_mar         = factor(sub_mar, levels = 0:1, labels = yesno),
    sub_opi         = factor(sub_opi, levels = 0:1, labels = yesno),
    sub_stim        = factor(sub_stim, levels = 0:1, labels = yesno),
    sub_herb        = factor(sub_herb, levels = 0:1, labels = yesno),
    sub_other       = factor(sub_other, levels = 0:1, labels = yesno),
    sub_none        = factor(sub_none, levels = 0:1, labels = yesno),
    
    num_drinks      = factor(num_drinks1, levels = drinks, labels = drinks),
    subs_time       = factor(subs_time, levels = time, labels = time),
    first_drink     = factor(first_drink, levels = time, labels = time),
    last_drink      = factor(last_drink, levels = time, labels = time),
    num_cigs        = factor(num_cigs1, levels = cig_count, labels = cig_count),
    first_cig       = factor(first_cig, levels = cigs_time, labels = cigs_time),
    last_cig        = factor(last_cig, levels = time, labels = time),

    op_heroin       = factor(op_heroin, levels = 0:1, labels = yesno),
    op_methadone    = factor(op_methadone, levels = 0:1, labels = yesno),
    op_pills        = factor(op_pills, levels = 0:1, labels = yesno),
    op_other        = factor(op_other, levels = 0:1, labels = yesno),
    num_opiate      = factor(num_opiate, levels = how_many, labels = how_many),
    
    stim_coc        = factor(stim_coc, levels = 0:1, labels = yesno),
    stim_meth       = factor(stim_meth, levels = 0:1, labels = yesno),
    stim_pills      = factor(stim_pills, levels = 0:1, labels = yesno),
    stim_other      = factor(stim_other, levels = 0:1, labels = yesno)
  )

# Clean environment
rm(list = setdiff(ls(), "daily"))

# Data check
about_data(daily)
# 2016-07-22: 7,378 observations and 172 variables in the data

test_that("The number of variables hasn't changed", {
  expect_equal(ncol(daily), 172)
})
```



-------------------------------------------------------------------------------   

&nbsp;

# Rename selected variables

```{r rename}
daily <- daily %>%
  rename(
    min_talk_c  = min_talk2,
    min_group_c = min_group2,
    num_cigs_c  = num_cigs2
  )
```


-------------------------------------------------------------------------------   

&nbsp;

## Select variables for analysis data

```{r select_vars}
daily <- daily %>%
  select(
    case_number,
    date,
    happy,
    frustrated,
    sad,
    worried,
    restless,
    excited,
    calm,
    lonely,
    bored,
    sluggish,
    loc_bus,
    loc_church,
    loc_home,
    loc_frifam,
    loc_store,
    loc_outdoors,
    loc_other,
    loc_restaurant,
    loc_school,
    loc_work,
    act_other,
    act_sit,
    act_sleep,
    act_stand,
    act_walk,
    act_talk,
    pre_who,
    yest_bike,
    yest_run,
    yest_cleaning,
    yest_none,
    min_walk,
    min_walk_c,
    min_run,
    min_run_c,
    min_act,
    min_act_c,
    min_sit,
    min_sit_c,
    pedometer,
    steps,
    phys_act_sat,
    serv_fruit,
    any_fruit,
    serv_veg,
    any_veg,
    serv_ssb,
    any_ssb,
    serv_sweets,
    any_sweets,
    serv_meat,
    any_meat,
    food_sat,
    take_meds,
    medtype_dep,
    medtype_psy,
    medtype_ast,
    medtype_bp,
    medtype_diab,
    medtype_pain,
    medtype_chol,
    medtype_anti,
    medtype_other,
    meds_ran_out,
    meds_forgot,
    meds_no_need,
    meds_side,
    meds_lost,
    meds_other,
    free_time,
    free_none,
    free_tv,
    free_games,
    free_movie,
    free_internet,
    free_book,
    free_music,
    free_craft,
    free_cleaned,
    free_pets,
    free_pray,
    free_sport,
    free_walk,
    free_cards,
    free_shop,
    free_volunteer,
    free_social,
    free_time_sat,
    min_talk,
    any_talk,
    min_group,
    any_group,
    interaction_sat,
    sub_alc,
    sub_cig,
    sub_mar,
    sub_opi,
    sub_stim,
    sub_herb,
    sub_other,
    sub_none,
    num_drinks,
    heavy_drink,
    subs_time,
    first_drink,
    last_drink,
    num_cigs,
    first_cig,
    last_cig,
    op_heroin,
    op_methadone,
    op_pills,
    op_other,
    num_opiate,
    stim_coc,
    stim_meth,
    stim_pills,
    stim_other,
    num_coc,
    num_stim,
    initial_alert_time,
    start_time,
    end_time,
    gps,
    dob,
    gender,
    hispanic,
    race,
    race_other
  )

# Data check
about_data(daily)
# 2016-07-22: 7,378 observations and 131 variables in the data

test_that("The number of variables hasn't changed", {
  expect_equal(ncol(daily), 131)
})
```


-------------------------------------------------------------------------------   

&nbsp; 

## Create lag variables

* The data is currently set up such that emotion today is paired with activity / substance use yesterday. For example, participant 10286 was asked about happiness on April 1 and about substance use yesterday (March 31st) on April 1.

* If our goal is to use emotional state today to predict activity / substance use _today_, then we need to create leading emotion variables or a lagged outcome variables.

* Further, participants don't respond (data isn't observed) everyday. So, we can't uniformly lag (or lead) by one day. We need to create a more intelligent lagged variable.

    + Create a variable (date_yest) that is equal to yesterday's date.
    
    + Create a variable (date_lag) that is equal to the date value for the preceding row, within person.
    
    + Create a dummy variable (date\_match) that is equal to 1 when date\_yest and date_lag are equal.
    
    + When date_match equals 1, then the lagged emotion value will be the emotion reported on the day that activity / substance use was reported. For example, participant 10286 was asked about happiness on April 1 and about substance use yesterday (April 1) on April 2. The lagged value of for happiness (April 1) in the April 2nd row and the value for substance use yesterday (April 1) in the April 2nd row, gives us the combination of emotion and outcome we are interested in.
    
* The same idea applies to location at time of response and activity at time of response.
    
```{r lagged_vars}
# Sort by case_number and date
daily <- arrange(daily, case_number, date)

# Create date variables
daily <- daily %>%
  group_by(case_number) %>%
  mutate(
    date_yest = as.Date(as.character(date - 1)),
    date_lag  = as.Date(as.character(lag(date)))
  ) %>%
  ungroup

# Create date match variable
daily <- daily %>% 
  mutate(
    date_match = if_else(as.numeric(date_yest) == as.numeric(date_lag), 1, 0)
  )

# Data check
# View(select(daily, date, date_yest, date_lag, date_match))

# Create lagged mood, location, activity, interacting variables
daily <- daily %>%
  group_by(case_number) %>%
  mutate_at(
    vars(happy:pre_who),
    funs("lag" = lag)) %>%
  ungroup

# Data check
# View(select(daily, date, date_yest, date_lag, date_match, happy, happy_lag))

# Move "lag"" suffix to prefix
pattern <- paste0("^(.*)_(", paste("lag", collapse = "|"), ")$")
daily <- daily %>% setNames(sub(pattern, "\\2_\\1", names(.)))
rm(pattern)

# Data check:
about_data(daily)
# 2016-07-22: 7,378 observations and 161 variables in the data

test_that("The number of variables hasn't changed", {
  expect_equal(ncol(daily), 161)
})
```

-------------------------------------------------------------------------------   

&nbsp; 

## Coding for CMOA plot

```{r cmoa_data}
# Happy
# X ranges from 0 to 4
# Y is fixed at 0
happy_axis <- function(x) {
  new <- NA
  new[x == "Strongly disagree"] <- 0
  new[x == "Disagree"] <- 1
  new[x == "Neutral"] <- 2
  new[x == "Agree"] <- 3
  new[x == "Strongly agree"] <- 4
  return(new)
}

# Sad
# X ranges from 0 to -4
# Y is fixed at 0
sad_axis <- function(x) {
  new <- NA
  new[x == "Strongly disagree"] <- 0
  new[x == "Disagree"] <- -1
  new[x == "Neutral"] <- -2
  new[x == "Agree"] <- -3
  new[x == "Strongly agree"] <- -4
  return(new)
}

# For emotions in quadrant 1
# X and Y range from 0 to 4
quad1 <- function(x) {
  new <- NA
  new[x == "Strongly disagree"] <- 0
  new[x == "Disagree"] <- 1
  new[x == "Neutral"] <- 2
  new[x == "Agree"] <- 3
  new[x == "Strongly agree"] <- 4
  return(new)
}

# For emotions in quadrant 2
# X ranges from 0 to -4
# Y ranges from 0 to 4
quad2x <- function(x) {
  new <- NA
  new[x == "Strongly disagree"] <- 0
  new[x == "Disagree"] <- -1
  new[x == "Neutral"] <- -2
  new[x == "Agree"] <- -3
  new[x == "Strongly agree"] <- -4
  return(new)
}

quad2y <- function(x) {
  new <- NA
  new[x == "Strongly disagree"] <- 0
  new[x == "Disagree"] <- 1
  new[x == "Neutral"] <- 2
  new[x == "Agree"] <- 3
  new[x == "Strongly agree"] <- 4
  return(new)
}

# For emotions in quadrant 3
# X and Y range from 0 to -4
quad3 <- function(x) {
  new <- NA
  new[x == "Strongly disagree"] <- 0
  new[x == "Disagree"] <- -1
  new[x == "Neutral"] <- -2
  new[x == "Agree"] <- -3
  new[x == "Strongly agree"] <- -4
  return(new)
}

# For emotions in quadrant 4
# X ranges from 0 to 4
# Y ranges from 0 to -4
quad4x <- function(x) {
  new <- NA
  new[x == "Strongly disagree"] <- 0
  new[x == "Disagree"] <- 1
  new[x == "Neutral"] <- 2
  new[x == "Agree"] <- 3
  new[x == "Strongly agree"] <- 4
  return(new)
}

quad4y <- function(x) {
  new <- NA
  new[x == "Strongly disagree"] <- 0
  new[x == "Disagree"] <- -1
  new[x == "Neutral"] <- -2
  new[x == "Agree"] <- -3
  new[x == "Strongly agree"] <- -4
  return(new)
}

# Create x and y coordinates for each row
daily <- daily %>%
  mutate(
    x_happy      = ifelse(date_match == 1, happy_axis(lag_happy), NA),
    y_happy      = 0,
    x_frustrated = ifelse(date_match == 1, quad2x(lag_frustrated), NA),
    y_frustrated = ifelse(date_match == 1, quad2y(lag_frustrated), NA),
    x_sad        = ifelse(date_match == 1, sad_axis(lag_sad), NA),
    y_sad        = 0,
    x_worried    = ifelse(date_match == 1, quad2x(lag_worried), NA),
    y_worried    = ifelse(date_match == 1, quad2y(lag_worried), NA),
    x_restless   = ifelse(date_match == 1, quad2x(lag_restless), NA),
    y_restless   = ifelse(date_match == 1, quad2y(lag_restless), NA),
    x_excited    = ifelse(date_match == 1, quad1(lag_excited), NA),
    y_excited    = ifelse(date_match == 1, quad1(lag_excited), NA),
    x_calm       = ifelse(date_match == 1, quad4x(lag_calm), NA),
    y_calm       = ifelse(date_match == 1, quad4y(lag_calm), NA),
    x_lonely     = ifelse(date_match == 1, quad3(lag_lonely), NA),
    y_lonely     = ifelse(date_match == 1, quad3(lag_lonely), NA),
    x_bored      = ifelse(date_match == 1, quad3(lag_bored), NA),
    y_bored      = ifelse(date_match == 1, quad3(lag_bored), NA),
    x_sluggish   = ifelse(date_match == 1, quad3(lag_sluggish), NA),
    y_sluggish   = ifelse(date_match == 1, quad3(lag_sluggish), NA)
  )

# Aggregate across all X values
# Not currently using frustrated, worried, lonely, or sluggish
daily <- daily %>%
  mutate(
    x = x_happy + x_sad + x_restless + x_excited + x_calm + x_bored,
    y = y_happy + y_sad + y_restless + y_excited + y_calm + y_bored
  )

rm(list = setdiff(ls(), "daily"))

# Data check:
# View(daily[c("case_number", "date", "date_yest", "lag_date", "date_match", "sub_alc", "happy", "lag_happy", "x_happy", "y_happy", "x_sad", "x_restless", "x_excited", "x_calm", "x_bored", "x")])

about_data(daily)
# 2016-07-22: 7,378 observations and 183 variables in the data

test_that("The number of variables hasn't changed", {
  expect_equal(ncol(daily), 183)
})
```


-------------------------------------------------------------------------------

&nbsp;

## Save as RData
```{r save_rdata}
about_data(daily)
save(daily, file = "data/daily_ema.RData")
```

**Save a .csv**
```{r save_csv}
write_csv(daily, path = "data/daily_ema.csv")
```


**Clean up environment**
```{r cleanup}
rm(list = setdiff(ls(), "daily"))
```

-------------------------------------------------------------------------------

&nbsp;

#### Session Info:
```{r session info, echo=FALSE}
sessionInfo()
```
